---
title: DANL Project
subtitle: "Data-Driven Mastery: Unlocking Business Potential"
author: 
  - __Daniel Noone__
  - __Angelina Batista__
  - __Elizabeth Dutan__
  - __Edward Eforo__
  - __Jaxson Marsh__

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```



# Introduction

The data frame being analyzed concerns sleep health data relating to different aspects of lifestyle. The data was found the website Kaggle found [here](https://www.kaggle.com/datasets/uom190346a/sleep-health-and-lifestyle-dataset).

## Description of Variables

- `Person ID`: Unique identifier for each individual

- `Gender`: Gender of the person

- `Age`: The age of the person in years

- `Occupation`: Occupation of the person

- `Sleep Duration`: Length of time person sleeps per day (in hours)

- `Quality of Sleep`: Subjective rating from 1-10 of quality of sleep

- `Physical Activity`: Length of time person engages in physical activity per day (minutes per day)

- `Stress Level`: Subjective rating of stress level experienced by person from 1-10

- `BMI Category`: BMI category of the person
    - Underweight
    - Normal
    - Overweight
  
- `Blood Pressure`: Blood pressure measurement of the person, systolic over diastolic pressure

- `Heart Rate`: The resting heart rate of the person in beats per minute

- `Daily Steps`: The number of steps person takes per day

- `Sleep Disorder`: The presence or absence of a sleep disorder in the person
    - None
    - Insomnia
    - Sleep Apnea


## Summary Statistics

```{r}
sleephlth <- read.csv("~/SUNY Geneseo/Classes/Fall 2023/DANL 200 (Team Pr)/dan-noone.github.io/Sleep_health_and_lifestyle_dataset.csv")
```

```{r}
sleephlth_sum <- skim(sleephlth)
sleephlth_sum
```




## MPG and a Type of Cars
The following boxplot shows how the distribution of highway MPG (`hwy`) varies by a type of cars (`class`) :blue_car: :truck: :minibus:.


```{r}
ggplot(data = mpg) +
  geom_boxplot(aes(x = class, y = hwy, fill = class),
               show.legend = F) +
  labs(x = "Class", y = "Highway\nMPG") 
```


